# Waseca Soil Pits

date: 2020-07-11

```{r setup-waseca, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(readxl)
library(aqp)
library(munsell)
library(soilDB)
```


## About 

Soil pits in Vivian Township, Waseca County, MN. Just north of Faribault county line.  Visited July 8, 2020 for filming of UMN Extension educational video on soil structure.  

Pea field with 2 pits dug by cooperator.  First pit in the headlands where pea health was very poor.  Second pit directly north ~30m from headland pit.

Field pit slightly upland from headland pit.  Cattails visible in field corner not far from headlands pit - clearly a wet area.  Field tiled for drainage, tile ~3 feet deep in this area.

**Headlands Pit** 
Canisteo?  Official Series Description [here](https://soilseries.sc.egov.usda.gov/OSD_Docs/C/CANISTEO.html)

Located at 43.8483, -93.6614

***Field Pit*** 
Nicollet? Official Series Description [here](https://soilseries.sc.egov.usda.gov/OSD_Docs/N/NICOLLET.html)

Located at 43.8487, -93.6614

## Description data collected

```{r data-waseca, echo=F}
data_raw <- read_excel("./data/waseca_pits.xlsx")

# aqp requires dataframe object
data_df <- as.data.frame(data_raw)

data_df <- data_df %>% 
  rename(id = Series, top = Depth_Top, bottom = Depth_Bottom, hzname = Horizon, hue = MoistHue1, value = MoistValue1, chroma = MoistChroma1)

#convert munsell colors to RGB
data_pits <- data_df %>% 
  mutate(soil_color = munsell2rgb(hue, value, chroma))

data_pits[ ,1:11]

data_pits %>% 
  select(PedonID, id, hzname, Redox)

data_pits %>% 
  select(PedonID, id, hzname, Notes)

```

## Create SPC objects

SPC = SoilProfileCollection 

```{r spc-obj-waseca}
# promote data_min_st dataframe to SoilProfileCollection 
depths(data_pits) <- id ~ top + bottom

#check that we have successfully converted class to SPC
class(data_pits)

# OSD data for the two series I think we have at the Waseca pits
osd_pedons <- fetchOSD(c('canisteo', 'nicollet'))

```

## Compare field descriptions with OSDs
```{r pedon-plot}
# join the OSD pedons with the pit pedons
both <- aqp::union(list(osd_pedons, data_pits))

# set margins
par(mar = c(5,3,2,2), xpd=NA)

# plot soil profile collection
plotSPC(both, width = 0.25, name = 'hzname', plot.order= c(1,3,2,4), cex.names = 0.7)

```

