# Mower County Pedons

date: 2020-09-06

## Overview

Visualizing some soil profile descriptions from SL in Mower County.  

This SPC plotting ideas post was helpful but I had to dig in the documentation to find it - not linked on the main aqp webpage? 
http://ncss-tech.github.io/AQP/aqp/SPC-plotting-ideas.html 

```{r setup-mw, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(tidy = TRUE)
knitr::opts_chunk$set(tidy.opts = list(width.cutoff=60))

library(tidyverse)
library(aqp)
library(readxl)
library(munsell)

```


## Load data

```{r data-mw}

pedons <- read_excel("./data/mower_pedons.xlsx") %>% 
  as.data.frame()

sites <- read_excel("./data/mower_site.xlsx")

```

## Munsell colors conversion

```{r}
#convert munsell colors to R compatible colors and add horizon ID
with_colors <- pedons %>% 
  mutate(soil_color = munsell2rgb(hue, value, chroma),
         hzID = c(1:nrow(pedons)))

with_colors

```

## Promote dataframe to SPC

SPC = soil profile collection (S4 object)

```{r}
# promote dataframe to SPC object 
depths(with_colors) <- id ~ top + bottom

# should be "SoilProfileCollection"
class(with_colors)

# inspect output 
str(with_colors)

# change the depth units (metadata/leabel) to inches - default is cm
depth_units(with_colors) <- 'inches'

# check that unit conversion worked
metadata(with_colors)

```

## Plot the SPC object

Most basic version here.

```{r}
# margin specification (bottom, left, top, right)
# default is typically c(5,4,4,2)
par(mar = c(1,1,1,1))

plot(with_colors, name = "name", width = 0.2)

```

## Add dashed lines 

Want to represent the lag line (transition to older till parent material) with a dotted line across each soil profile.

```{r}
# grab lag line depth from sites df

lag <- sites %>% 
  select(id, lag_in)

# need the ids in alpha order to align with pedons plotted alphabetically below,
# (otherwise the lag lines get plotted on the wrong pedon).  
# Find a more robust solution for this in the future

lag_sorted <- lag[order(lag$id),]

# keep in mind that each pedon is centered over its integer index on the x-axis of the plot (first pedon centered over 1, second over 2, etc.)
x.pos <- 1:length(with_colors)

# segments function needs vectors of coordinates:
#specifies start/end the line segments
# see https://bookdown.org/ndphillips/YaRrr/low-level-plotting-functions.html
from.x <- c(x.pos - 0.2)
to.x <- c(x.pos + 0.2)
from.y <- lag_sorted$lag_in
to.y <- lag_sorted$lag_in 

par(mar = c(0,2,0,2))

plot(with_colors, name = "name",
     width = 0.2)
# lwd is line width, lty is linetype
segments(x0 = from.x,
         x1 = to.x,
         y0 = from.y,
         y1 = to.y, 
         col = "red", 
         lwd = 3,
         lty = 3)


```

## References 

Low level plotting functions, including segments() and text(), which might be useful for adding additional labels (such as "lag line") in the future:  https://bookdown.org/ndphillips/YaRrr/low-level-plotting-functions.html  

SPC plotting ideas: 
http://ncss-tech.github.io/AQP/aqp/SPC-plotting-ideas.html  



```{r echo=TRUE, include=TRUE, results='markup'}
devtools::session_info()
```

